On Similarity

### 2016-05-03 paper.tex
Sentences

In the previous section we illustrated similarity with examples between word pairs.

>>>>

In the previous section we focused on the properties of similarity, illustrating it with word-word examples. This section is dedicated to the nuances of similarity measurement between words, phrases and sentences. We assume that similarity requires the agreement of the grammatical roles of the words or phrases.

A noun phrase can be similar to a noun, as in \textit{female lion} and \textit{lioness}, as between noun phrases with modifiers (\textit{yellow car} and \textit{taxi}), so the same similarity principles can be applied to phrases as to words. In this case, similarity is measured in \emph{context}. It is still comparison of head terms which meaning is modified by arguments they appear with \cite{Dinu:2010:MDS:1870658.1870771,mitchell2010composition}. With verbs this idea can be applied to compare transitive verbs with
intransitive. For example, \textit{cycle} is similar to \textit{ride a bicycle}.

Sentential similarity can be explained in several ways. First of all, it might be seen as the similarity of the heads in the contexts. That is, the similarity between \textit{sees} and \textit{notices} in \textit{John \textbf{sees} Mary} and \textit{John \textbf{notices} a woman}. This approach abstracts away grammatical difference between the sentences and concentrates on semantics and fits the proposed model as the respect for the head, which is a lexical entity, has to be found.

Sentences are different from words because they are part of the discourse and it is easy to come up with a pair of sentences that are similar in a certain situation but don not share grammatical structure at all, for example \textit{Yes} and \textit{They are on the table} are similar sentences if they are answers to the question \textit{Where are the cookies?} Questions the sentences answer are valid respects for similarity measurements, as well as entailment, paraphrase \cite{White:2015:WSE:2838931.2838932} or spatial categories \cite{ritter-EtAl:2015:*SEM2015}.
